On the Bit Complexity of Finding Points in Connected Components of a Smooth Real Hypersurface

### 2020-02-02 main.tex
Update main.tex
Updated Section 6

It will sometimes be useful for us to let the matrix $A$ represent new
indeterminates: let $\A$ be an $n \times n$ matrix of new
indeterminates $(\A_{i,j})_{1 \leq i , j \leq n}.$ Define $f^{\A} \in
\mathbb{Q}(\A_{i,j})[\textbf{X}]$ as $f(\A\textbf{X})$ and
$V^\A=V(f^\A).$ It follows from Proposition 4.2 that for a generic $A
\in \C^{n^2},$ $0$ is a regular value of $\Phi_{A}$.  The proof of
\cite[Theorem B.3]{NO}, which shows the existence of $\Delta \subset
\C^{n^2},$ applies \textit{Sard's lemma}, which shows that the
critical values of a polynomial mapping are contained in a
hypersurface. It follows from Sard's lemma that for the matrix of
indeterminants $\A,$ $0$ is a regular value of $\Phi_{\A}.$ Otherwise
it could not be that $0$ is a regular value of $\Phi_A,$ for a generic
$A\in \C^{n^2}.$ Therefore,

@@

\jac_{\xb}(\Phi_{\A}) = \jac_{\xb}\left(f^{\A}(x),\frac{\pa f^{\A}(x)}{\pa X_{i+1}},\hdots,\frac{\pa f^{\A}(x)}{\pa X_n}\right)

@@

has full rank for all $\xb \in V(f^{\A})$, and thus $f^{\A}$satisfies $\textbf{H}_i'(1),\textbf{H}_i'(2)$ and $\textbf{H}_i'(3)$. We will rely on this fact in Section 5. 

@@

We assume our input polynomial $f$ satisfies \textbf{H} so that $f$ is squarefree with $V(f) \cap \mathbb{R}^n$ smooth. Thus, by Proposition 2.4,  $W(\pi_i,V(f))$ is defined by the vanishing of \[\fb := \left(f,\frac{\pa f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n}\right).\]
If the Jacobian of $\fb$ has full rank at all $x \in V(f) \cap \mathbb{R}^n,$ then it follows by the Jacobian criterion [4, Theorem 16.19] that $f$ satisfies $\textbf{H}_i'(1),\textbf{H}_i'(2)$ and $\textbf{H}_i'(3)$. 

@@

There exists a Zariski closed set $\Delta \subset \C^{n^2}$ of degree at most $D^n$ with the property that, if $A \in \C^{n^2} - \Delta$ then for each $i \in \{1,\hdots,n\},$ $f^A$ satisfies $\textbf{H}_i'(1),\textbf{H}_i'(2)$ and $\textbf{H}_i'(3)$.

@@

Then we show that $0$ is a regular value of $\Phi$, so that, by Proposition 4.2, $\Delta \subset \C^{n^2}$ exists with the property that if $A \in \C^{n^2}-\Delta$ then $0$ is a regular value of $\Phi_A.$ It then follows that 

@@

has full rank for all $\xb \in V(f^A)$. Then, by the Jacobian criterion \cite[Theorem 16.19]{ECA}, $f^A$ satisfies $\textbf{H}_i'(1),\textbf{H}_i'(2)$ and $\textbf{H}_i'(3)$.

@@

Now, with $i \in \{1,\hdots,n\}$, let $\Phi$ be the mapping \[(\xb,A) \mapsto (f^A(\xb), \frac{\partial f^A}{\partial X_{i+1}}(\xb),\hdots,
\frac{\partial f^A}{\partial X_n}(\xb))\] so that 
\Phi^{-1}(0)= W(\pi_i,V^A).

@@

Put

@@

G_0 = f(Ax), 
G_1 = \frac{\partial (f(Ax))}{\partial X_{i+1}},\hdots,G_{n-i} = \frac{\partial (f(Ax))}{\partial X_n},

@@

and assume that $G_0(\xb,A)=G_1(\xb,A)=\hdots =G_{n-i}(\xb,A)=0.$ We can simplify by removing $Ax$ from the system because  

@@

(f(Ax),(\grad(f)\cdot 
\bbm 
a_{i+1,1} \\
\vdots \\
a_{i+1,n} 
\ebm)(Ax),\hdots, 
\bbm 
a_{n,1} \\
\vdots \\
a_{n,n} 
\ebm)(Ax)

@@

defines a radical ideal of dimension ? whose variety is smooth if and only if 
(f,\grad(f)\cdot 
\bbm 
a_{i+1,1} \\
\vdots \\
a_{i+1,n} 
\ebm),\hdots, 
\bbm 
a_{n1} \\
\vdots \\
a_{nn} 
\ebm)
defines a radical ideal of dimension ? whose variety is smooth. Now, let $g$ be the following:

@@

    g: \C^n \times \C^{n^2} &\rightarrow \C^{n-i+1} \\

@@

@@

\right)

@@

I claim that for all $(\xb,A) \in g^{-1}(0),$ the Jacobian matrix of $g$ at $(\xb,A)$ has full rank $n-i+1$ (i.e. $g$ is transverse to $\{0\}).$ The Jacobian of 
       & \left(f,\grad(f)\cdot 
\bbm 
a_{i+1,1} \\
\vdots \\
a_{i+1,n} 
\ebm,\hdots, 
\bbm 
a_{n,1} \\
\vdots \\
a_{n,n} 
\ebm\right) \\
&= \left(f, \frac{\pa f}{\pa X_1} a_{i+1,1} + \hdots + \frac{\pa f}{\pa X_n}a_{i+1,n}, \hdots, \frac{\pa f}{\pa X_1} a_{n,1} + \hdots + \frac{\pa f}{\pa X_n}a_{n,n}\right) \\
&= (f,F_{1},\hdots,F_{n-i}),  
where 

@@

F_j = \frac{\pa f}{\pa X_1} a_{i+j,1} + \hdots + \frac{\pa f}{\pa X_n}a_{i+j,n}, 1 \leq j \leq n-i,\]
is

@@

Since $f$ is square-free, \[I(V(f)) = \langle f \rangle,\]  and thus for all $\xb \in V(f)$, 

@@

Since $\deg \Phi \leq D,$ it follows by Theorem 4.2 that a hypersurface $\Delta \subset \C^{n^2}$ exists, of degree at most $D^n,$ with the property that, if $A \in \C^{n^2}-\Delta,$ then $0$ is a regular value of $\Phi_{A}$, which means that 

@@

has full rank for all $\xb \in V(f^A)$ ($\cap \R^n$?). Therefore, by the Jacobian criterion \cite[Theorem 16.19]{ECA}, $f^A$ satisfies $\textbf{H}_i'(1),\textbf{H}_i'(2)$ and $\textbf{H}_i'(3)$, for each $i \in \{1,\hdots,n\}.$

>>>>

It will sometimes be useful for us use in replace of the matrix $A$, a matrix representing indeterminants: let $\A$ be an $n \times n$ matrix of new indeterminates $(\A_{i,j})_{1 \leq i , j \leq n}.$ Define $f^{\A} \in \mathbb{Q}(\A_{i,j})[\textbf{X}]$ as $f(\A\textbf{X})$ and $V^\A=V(f^\A).$ It follows from Proposition 5.1, that for a generic $A \in \C^{n^2},$ $0$ is a regular value of $\Phi_{A}$.  The proof of \cite[Theorem B.3]{TWT}, which establishes the existence of $\Delta \subset \C^{n^2},$ applies \textit{Sard's lemma}, which shows that the critical values of a polynomial mapping are contained in a hypersurface. It follows from Sard's lemma that, for the matrix of indeterminants $\A,$ $0$ is a regular value of $\Phi_{\A}.$ Otherwise it could not be that $0$ is a regular value of $\Phi_A,$ for a generic $A\in \C^{n^2}.$ The proof is routine and we omit details. It then follows that 

@@

\jac_{\xb}(\Phi_{\A}) = \jac_{\xb}\left(f^{\A}(\xb),\frac{\pa f^{\A}(\xb)}{\pa X_{i+1}},\hdots,\frac{\pa f^{\A}(\xb)}{\pa X_n}\right)

@@

has full rank for all $\xb \in V(f^{\A})$, and thus $f^{\A}$satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$ by the Jacobian criterion \cite[Theorem 16.19]{ECA}. We will rely on this fact in Section 8.

@@

We assume our input polynomial $f$ satisfies \textbf{H} so that $f$ is squarefree with $V(f) \cap \mathbb{R}^n$ smooth. Thus, by Proposition 2.4,  $W(\pi_i,V(f))$ is defined by the vanishing of $f,\frac{\pa f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n}.$ If the Jacobian of this set of polynomials has full rank at all $\xb \in V(f) \cap \mathbb{R}^n,$ then it follows by the Jacobian criterion \cite[Theorem 16.19]{ECA} that $f$ satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$. 

@@

There exists a Zariski closed set $\Delta_1 \subset \C^{n^2}$ of degree at most $D^n$ with the property that, if $A \in \C^{n^2} - \Delta_1$ then for each $i \in \{1,\hdots,n\},$ $f^A$ satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$.

@@

We want to apply Theorem 5.1, and thus we show that $0$ is a regular value of $\Phi$, where then it follows that $\Delta_1 \subset \C^{n^2}$ exists with the property that if $A \in \C^{n^2}-\Delta_1$ then $0$ is a regular value of $\Phi_A,$ which means that 

@@

has full rank for all $\xb \in V(f^A)$. Then, by the Jacobian criterion \cite[Theorem 16.19]{ECA}, $f^A$ satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$.

@@

With $i \in \{1,\hdots,n\}$, let $\Phi$ be as stated above.

@@

Let $A_1,\hdots,A_n$ denote the columns of $A,$ and let $(\xb,A) \in \C^n \times \C^{n^2}$ be such that \[
f^A(\xb) = \frac{\partial f^A(\xb)}{\partial X_{i+1}}=\hdots= \frac{\partial f^A(\xb)}{\partial X_n}=0.
We want to show that the system

@@

f(Ax),(\grad(f)\cdot 
A_{i+1})(Ax),\hdots, 
A_n)(Ax)

@@

defines a radical ideal of dimension ? whose variety is smooth, which holds if and only if 

@@

f,\grad(f)\cdot 
A_n

@@

defines a radical ideal of dimension ? whose variety is smooth. Now, let $\Psi$ be the mapping

@@

    \Psi: \C^n \times \C^{n^2} &\rightarrow \C^{n-i+1} \\

@@

A_{i+1},\hdots. 

@@

\right),

@@

For all $(\xb,A) \in \Psi^{-1}(0),$ the Jacobian matrix of $\Psi$ at $(\xb,A)$ has full rank $n-i+1$. Indeed, set 

@@

F_j = \frac{\pa f}{\pa X_1} a_{i+j,1} + \hdots + \frac{\pa f}{\pa X_n}a_{i+j,n}, ~1 \leq j \leq n-i,
The Jacobian of the system of polynomials $f,F_{1},\hdots,F_{n-i}$ is

@@

Since $f$ is square-free, $I(V(f)) = \langle f \rangle.$ Therefore, for all $\xb \in V(f)$, 

@@

Since $\deg \Phi \leq D,$ it follows by Theorem 5.1 that a hypersurface $\Delta_1 \subset \C^{n^2}$ exists, of degree at most $D^n,$ with the property that, if $A \in \C^{n^2}-\Delta_1,$ then $0$ is a regular value of $\Phi_{A}$, which means that 

@@

has full rank for all $\xb \in V(f^A)$ ($\cap \R^n$?). Therefore, for each $i \in \{1,\hdots,n\},$ by the Jacobian criterion \cite[Theorem 16.19]{ECA}, $f^A$ satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$.%
